---
title: "Young People Survey"
author: "Darian-Florian Voda"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
library(ggplot2)
library(dplyr)
library(ggrepel)
library(tidyverse)
```

## R Data Visualization Project: Young People Survey

## Information about the main dataset (responses)

- In 2013, students of the Statistics class at FSEV UK were asked to invite their friends to participate in this survey.
- The data file (responses.csv) consists of 1010 rows and 150 columns (139 integer and 11 categorical).
- For convenience, the original variable names were shortened in the data file.
- File columns.csv represents the dataset for the shortenings made for each variable name
- All participants were of Slovakian nationality (the questionnaire was later translated in English), aged between 15-30.

The variables can be split into the following groups:
- Music preferences (19 items)
- Movie preferences (12 items)
- Hobbies & interests (32 items)
- Phobias (10 items)
- Health habits (3 items)
- Personality traits, views on life, & opinions (57 items)
- Spending habits (7 items)
- Demographics (10 items)

## Main Question: How to effectively visualize a lot of variables in order to gain some meaningful insights from the data?

### Example of a specific category

#### MUSIC PREFERENCES

1. I enjoy listening to music.: Strongly disagree 1-2-3-4-5 Strongly agree (integer)
2. I prefer.: Slow paced music 1-2-3-4-5 Fast paced music (integer)
3. Dance, Disco, Funk: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
4. Folk music: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
5. Country: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
6. Classical: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
7. Musicals: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
8. Pop: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
9. Rock: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)
10. Metal, Hard rock: Don't enjoy at all 1-2-3-4-5 Enjoy very much (integer)



### Import data
```{r}
data = read.csv("responses.csv")
columns = read.csv("columns.csv")

head(data)
```
### See columns data

```{r}
head(columns)
```
## Exploratory Data Analysis (EDA)

Firstly, we will split each category in dataframes for a better management

```{r}
# Split into DFs
music_dataframe = data.frame(data[1:19])

movies_dataframe = data.frame(data[20:31])

hobbies_dataframe = data.frame(data[32:63])

phobias_dataframe = data.frame(data[64:73])

health_habits_dataframe = data.frame(data[74:76])

personality_dataframe = data.frame(data[78:133])

spendings_dataframe = data.frame(data[134:140])

demographics_dataframe = data.frame(data[141:150])
```

Now, we will take a look how much people enjoy the **Music** in general

Firstly create a dataframe of the frequencies for the specific question
```{r}
overall_music = music_dataframe %>% 
  group_by(Music) %>%
  summarise(counts = n())

overall_music
```

```{r}
overall_music_plot = ggplot(overall_music, aes(x = Music, y = counts, fill=counts)) +
  geom_bar(stat = "identity", fill="steelblue")+
  geom_text(aes(label= counts), vjust = -0.3)+
  theme(axis.title.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  labs(y="Number of respondents")+
  ggtitle("How much people enjoy Music?")+
  geom_text(x=1, y=770, label="1 - not so much", color="red", hjust = 0, size = 5)+
  geom_text(x=1, y=740, label="5 - very much", color = "blue", hjust = 0, size = 5)+
  scale_y_continuous(breaks=round(seq(0, max(overall_music$counts), by=100), 1))

overall_music_plot
```





## Idea:

- Create an an interactive graph that is based on the specific things (Music, Movies, etc)
- Then on the lower side of this mini-infograph you can expand the other things for the other categories
- Create Pie charts for Genres of music/movie/etc


