---
title: "Chapter 3"
author: "Darian-Florian Voda"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Data Visualization
### A practical Introduction
#### By Kieran Healy

```{r}
library(gapminder)
library(tidyverse)
library(ggplot2)
library(socviz)
gapminder
```

```{r}
p = ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line()
```

```{r}
p = ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(aes(group = country))
```

```{r}
p = ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(aes(group = country)) + facet_wrap(~continent)
```

```{r}
p = ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(color="gray70", aes(group = country)) +
  geom_smooth(size=1.1, method="loess", se=FALSE) +
  scale_y_log10(labels=scales::dollar) +
  facet_wrap(~continent, ncol=5) +
  labs(x="Year", y="GDP per capita", title="GDP per capita on Five Continents")
```

```{r}
glimpse(gss_sm)
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) +
    geom_smooth() + 
    facet_grid(sex ~ race)
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar()
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop..))
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop.., group = 1))
```

```{r}
table(gss_sm$religion)
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = religion, fill=religion))
p + geom_bar() + guides(fill = FALSE)
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion, fill=religion))
p + geom_bar()
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion, fill=religion))
p + geom_bar(position="fill")
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion, fill=religion))
p + geom_bar(position="dodge", mapping=aes(y = ..prop..))
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = bigregion, fill=religion))
p + geom_bar(position="dodge", mapping=aes(y = ..prop.., group=religion))
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = religion))
p + geom_bar(position="dodge", mapping=aes(y = ..prop.., group=bigregion)) + facet_wrap(~bigregion, ncol=1)
```

```{r}
p = ggplot(data = midwest, mapping = aes(x = area))
p + geom_histogram()
```

```{r}
p = ggplot(data = midwest, mapping = aes(x = area))
p + geom_histogram(bins= 10)
```

```{r}
oh_wi = c("OH", "WI")

p = ggplot(data = subset(midwest, subset = state %in% oh_wi),
           mapping = aes(x=percollege, fill = state))

p + geom_histogram(alpha=0.4, bins=20)
```

```{r}
p = ggplot(data = midwest, mapping = aes(x = area))
p + geom_density()
```

```{r}
p = ggplot(data = midwest, mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3)
```

```{r}
p = ggplot(data = subset(midwest, subset = state %in% oh_wi), mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3, mapping = (aes(y = ..scaled..)))
```

```{r}
titanic
```

```{r}
p = ggplot(data = titanic, mapping = aes(x = fate, y = percent, fill = sex))
p + geom_bar(position = "dodge", stat = "identity") + theme(legend.position="top")
```

```{r}
oecd_sum
```

```{r}
p = ggplot(data = oecd_sum, mapping = aes(x = year, y = diff, fill = hi_lo))
p + geom_col() + guides(fill=FALSE) + 
    labs(x = NULL, y = "Difference in Years",
         title = "The US Life Expectancy Gap",
         subtitle = "Difference between US and OECD average life expectancies, 1960-2015",
         caption = "Data: OECD. After a chart by Chritopher Ingraham, Washington Post, December 27th 2017.")
```

### 1. Gapminder plot

```{r}
library(here)
p = ggplot(data = gapminder, mapping = aes(x = pop, y = gdpPercap))
new_plot = p + geom_point(alpha = 0.2) +
    geom_smooth() + 
    facet_wrap(~year)

new_plot
ggsave(here("gapminder by year.pdf"), plot=new_plot)
```

### 2. facet_grid(sex~race)




```{r}
p = ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) +
    geom_smooth() + 
    facet_grid(sex ~ race)
```

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) +
    geom_smooth() + 
    facet_grid(~ sex + race)
```

The difference is that in the second facet grid we kind of making a filter based on sex + race as it was mentioned in this chapter because we have contiguous table.

### 3. facet_wrap(~ sex + race)

```{r}
p = ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) +
    geom_smooth() + 
    facet_wrap(~ sex + race)
```

There is an interesting difference betwee facet_grid and facet_wrap

### 4. Use of geom_freqpoly()

```{r}
p = ggplot(data = midwest, mapping = aes(x = area))
p + geom_freqpoly()
```

### 5. geom_bin2d()

```{r}
p = ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_bin2d(bins = c(20, 50))
```

UUuuu, I like that.

### 6. geom_density_2d()

```{r}
midwest
```


```{r}
p = ggplot(data = midwest, mapping = aes(x = percollege, y = percbelowpoverty))
p + geom_density_2d()
```

```{r}
p = ggplot(data = midwest, mapping = aes(x = percollege, y = percbelowpoverty))
p + geom_density_2d() + geom_point()
```

Very interesting chapter 4 and also interesting compulsory exercises!!
I have no questions for this chapter, only things that I newly learned.

With kind regards,
Darian